---
title: "Homework 4"
author: "Yuren Zhang"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: yes
    toc: true
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{homework 4 vignette}
-->

## 5.8.2

From CASL equation 2.73, 
$$cond(A)=\frac{\sigma_{max}}{\sigma_{min}}$$
$\sigma_{max}$: the largest singular value
$\sigma_{min}$: the smallest singular value

From CASL 2.77
```{r}
X <- matrix(c(1, 2, 2, 1), 2, 2)
p <- c(0.5, 0.0000001)
```

Then the condition number of $X'X$ is
```{r}
svals <- svd(t(X)%*%X)[["d"]]
max(svals)/min(svals)
```

From CASL 5.28
The Hessian matrix for Bernoulli random Variable is 
$$X'diag(p(1-p))X$$
The condition number of this Hessian matrix is 
```{r}
H = t(X)%*%diag(p*(1-p))%*%X
svals2 <- svd(H)[["d"]]
max(svals2)/min(svals2)
```

Comparing the Hessian Matrix of logistic variation $X'diag(p(1-p))X$ and linear Hessian $X'X$, we can see that the former one is much larger than the latter one. Thus the linear Hessian is well conditioned.

##5.8.4
From CASL p129

```{r}
# Solve generalized linear models with Newton-Ralphson method.
# Args:
# X: A numeric data matrix.
# y: Response vector.
# family: Instance of an R 'family' object.
# maxit: Integer maximum number of iterations.
# tol: Numeric tolerance parameter.
# Returns:Regression vector beta of length ncol(X).

irwls_glm <- function(X, y, family, maxit=25, tol=1e-10, lambda){
  beta <- rep(0,ncol(X))
  for(j in 1:maxit){
    b_old <- beta
    eta   <- X %*% beta
    mu    <- family$linkinv(eta)
    mu_p  <- family$mu.eta(eta)
    z     <- eta + (y - mu) / mu_p
    W     <- as.numeric(mu_p^2 / family$variance(mu))
    XtX   <- crossprod(X, diag(W) %*% X)
    Xtz   <- crossprod(X, W * z)
    V <- XtX + diag(lambda, dim(XtX)[1])
    beta  <- solve(V, Xtz)
    if(sqrt(crossprod(beta - b_old)) < tol) break
  }
  beta
}
```

##Q3

We first creat the sparse.matrix object:
```{r}
#' @description create sparse matrix using sparse.matrix, operations defined including add, multiply and transpose
#' @param i row index 
#' @param j col index 
#' @param x value of the (i,j) entry
#' @param dims sparse matrix dimension
#' @return A sparse.matrix object
#' @export

sparse.matrix = function(i, j, x, dims = c(max(i), max(j))){
  structure(list(data.frame(i = c(1, 2), j = c(1, 1), x = c(3, 1)), dims), class = "sparse.matrix")
}
```

an example will be:

```{r}
a = sparse.matrix(i = c(1, 2), j = c(1, 1), x = c(1, 1))
b = sparse.matrix(i = c(3, 2), j = c(2, 1), x = c(1, 4))
c = sparse.matrix(i = 1, j = 1, x = 1)
```

then addition function:
```{r}
#' @description create sparse matrix addition method.
#' @param a a sparse matrix
#' @param b a sparse matrix
#' @return A sparse.matrix object a + b
#' @export

sparse_add = function(a, b){
# making sure the format is correct
    if (!identical(a[[2]],b[[2]]))
    stop("non-conformable arrays")
  if (!inherits(a, "sparse.matrix"))
    stop ("a is not a sparse.matrix object")
  if (!inherits(b, "sparse.matrix"))
   stop ("b is not a sparse.matrix object")
  c = merge(a[[1]], b[[1]], by = c("i", "j"), all = TRUE, suffixes = c("1", "2"))
  c$x1[is.na(c$x1)] = 0
  c$x2[is.na(c$x2)] = 0
  c$x = c$x1 + c$x2
  c[, c("i", "j", "x")]
  c = sparse.matrix(c$i, c$j, c$x, dims = a[[2]])
  return(c)
}
```

an example will be
```{r}
sparse_add(a,a)
```

then multiply function:
```{r}
sparse_multi = function(a,b){
 if (!identical(a[[2]][2],b[[2]][1]))
    stop("non-conformable arguments")
  if (!inherits(a, "sparse.matrix"))
    stop ("a is not a sparse.matrix object")
  if (!inherits(b, "sparse.matrix"))
   stop ("b is not a sparse.matrix object")
  
#from CASL p294 to 295 casl_sparse_multiply we have
  
  colnames(b[[1]]) = c("i2", "j2", "x2")
  c = merge(a[[1]], b[[1]], by.x = "j", by.y = "i2",
             all = FALSE, suffixes = c("1", "2"))
  c$x = c$x * c$x2
  c$key = paste(c$i, c$j, sep = "-")
  x = tapply(c$x, c$key, sum)
  key = strsplit(names(x), "-")
  d = data.frame(i = sapply(key, getElement, 1),
                  j = sapply(key, getElement, 2),
                  x = as.numeric(x))
  d = sparse.matrix(c$i, c$j, c$x, dims = c(a[[2]][1], b[[2]][2]))
  return(d)
}
```

an example will be:
```{r}
sparse_multi(a,c)
```

last, transpose:
```{r}
#' @description create sparse matrix transpose method.
#' @param a a sparse matrix
#' @return A transposed sparse.matrix object t(x)
#' @export
sparse_t = function(a){
  temp = a[[1]]$i
  a[[1]]$i = a[[1]]$j
  a[[1]]$j = temp
  a[[2]] = rev(a[[2]])
  return(a)
}
```

an example will be:
```{r}
sparse_t(a)
```

to add all above functions into the sparse.matrix class, we need to overload but keep original algorithms. 
```{r}
#reference : http://adv-r.had.co.nz/S3.html

# %*% is not S3 object
`%*%.default` = .Primitive("%*%")  # keep defalut operation

`%*%` = function(x, y) {
  UseMethod("%*%", x)
}

t = function (x, ...) {
  UseMethod("t", x)}
```





